The role of advection versus diffusion in intraventricular transport

### 2025-02-05 paper/main.tex
fix figure and panel references

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask: what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}\fixme{e/x}). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the \fixme{different-characteristics}
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}a). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}\fixme{x-y}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site (\Cref{fig:fig2}\fixme{z}).
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of
    Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{b--d.} Slice of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for eight time instants covering the
    time range from 45 seconds to 790 seconds, corresponding to 100 to 1750 cardiac cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    The colorbar is scaled so the maximum is the maximum concentration of the Dendra2 
    condentration field slice at the last time instant.
    \textbf{e--g.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

To quantify the transport of solute, we plot the profile of the mean
concentrations $\cbar_i(t)$ as functions of time
(\Cref{fig:fig2}\fixme{b--d}) and
the times $\hat{t}$ when the mean concentrations $\cbar_i$ first exceed
a threshold value $\hat{c}$ (\Cref{fig:fig2}\fixme{f})
in 6 regions of interest (ROIs) (\Cref{fig:fig2}\fixme{e}).
The dynamics at the site of photoconversion (ROI 1)
is logarithmic and follows from imposing \eqref{eq:photoconversion_curve}.
For ROIs 2--6, we observe faster kinetics with higher diffusion coefficients
(\Cref{fig:fig2}\fixme{b--d}),
which results in smaller time to threshold
(\Cref{fig:fig2}\fixme{f}).
For all three diffusion coefficients, the dynamics of ROI 5
is significantly delayed compared to ROI 2--4,
and ROI 6 is the slowest. Notably, for $D_1$ the mean concentration
in ROI 6 is close to zero at the end of simulation.
To quantify the balance of advection versus diffusion as transport mehcanisms,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$,
$\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the
dominant transport mechanism on a \fixme{global} scale.

\subsection*{Validation of the simulations with experimental data}
To validate our computational model, we performed photoconversion experiments
in transgenic zebrafish larvae expressing a secreted Dendra2 in their brain ventricles.
Our photoconversion protocol consisted of the acquisition of a baseline,
before localized exposure to a UV laser (at the ROI 1 location)
that converts Dendra2 from a green to a red emitting fluorescent protein. 
To quantify the transport of photoconverted proteins, we measured and
calculated the change in fluorescence intensity $\Delta F = (F(t)-F_0)/F_0$
over time (\Cref{fig:figure3_compare_exp_and_sim_control}a,
$F_0$ being the average fluorescence at baseline).

We obtained very similar kinetics curves (\Cref{fig:figure3_compare_exp_and_sim_control}a, b)
and time-to-threshold values (\Cref{fig:figure3_compare_exp_and_sim_control}c, d) when comparing our
experiments (\Cref{fig:figure3_compare_exp_and_sim_control}a, c) with
simulation (\Cref{fig:figure3_compare_exp_and_sim_control}b, d).
Importantly, since the initial condition of the simulation is a mean concentration of zero,
the mean concentration $\cbar$ can be interpreted as the change in mean concentration and
is thus a similar metric to $\Delta F$.
The final values of $\cbar$ were 0.76, 0.72, 0.82, 0.56 and 0.35 respectively,
compared to 0.63, 0.71, 0.67, 0.29 and 0.18 for $\Delta F$.
Notwithstanding similarities, we observed small differences between the experiments and simulation.
For instance, transport dynamics in ROIs 2--6 were faster in the simulation and reached higher final values.
Moreover, the $\cbar$ profiles of ROIs 3--4 were opposite order in
the simulations as compared to the experiments.
Despite observing small differences, the \emph{in vivo} experiments and
simulations were similar enough to convincingly validate the model.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F$ over time
    in each region of interest (ROI)
    observed during experiments of photoconversion of Dendra2 proteins in ROI 1.
    \textbf{b.} Mean concentration $\cbar$ as function of time in each ROI in a
    simulation of Dendra2 protein transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F$ reaches a threshold value of 0.25 in
    one representative zebrafish larva.
    Time to $\cbar$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    We have excluded the posterior-most regions of the geometries,
    because the values of $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions.}
    \label{fig:figure3_compare_exp_and_sim_control}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.7637806  0.72025859 0.81903216 0.56204164 0.34926411

\subsection*{Absence of ciliary motion affects local protein distribution}
Since CSF flow is highly disturbed in the absence of motile cilia, we
want to assess the impact of cilia motion on solute distribution. To this end,
we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant zebrafish (\emph{schmalhans} (\emph{smh})).
By plotting the means of $\Delta F$ (\Cref{fig:figure4_compare_control_mutant}a)
and the time to threshold
across the control and mutant cohorts (\Cref{fig:figure4_compare_control_mutant}b),
we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in $\sim 45\%$ (ROI 2) and $\sim 25\%$ (ROI 4)
higher mean times-to-threshold in the \emph{smh}
mutants compared to controls (\Cref{fig:figure4_compare_control_mutant}b).

Next, to identify whether our model would recapitulate these findings,
we simulated protein transport by using the only-cardiac/no-cilia flow model 
and compared it with the baseline model
(both cilia and cardiac motion) (\Cref{fig:figure4_compare_control_mutant}c).
Similar to our experiments, we observed slower dynamics in simulations for ROIs 2 and 4.
Interestingly, the difference was even greater for the simulations,
with the time-to-threshold in ROI 2 being 90\% higher in absence of cilia-mediated flow.
Moreover, ROI 5 exhibits very different $\cbar$ profiles for the two
configurations in our computational model,
but not in our experimental data.
Altogether, our findings identify that diffusion plays an important
role for solute movement to regions upstream in the cilia flow fields or to distant ventricles.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F$ over
    time in each region of interest (ROI) for control (black lines) and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets, and $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%.
    \textbf{c.} Mean concentration $\cbar$ as function of time in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:figure4_compare_control_mutant}
\end{figure}

\subsection*{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we simulate transport
upon region-specific cilia paralysis by modifying the boundary conditions
of the CSF flow model. Three separate cilia-covered regions were considered separately:
the telencephalic (anterior) ventricle, the dorsal part and the ventral part
of the diencephalic (middle) ventricle (\Cref{fig:figure5_compare_cilia_modifications}a).
The cilia tangential traction boundary condition is for each case
changed to a free-slip boundary condition in paralyzed region,
prescribing zero tangential forces to
represent absence of cilia motion.
We simulate CSF flow modeled by the Stokes equations with the
modified boundary conditions, and use the resulting
velocity fields in transport simulations.

We identified that the velocities were most strongly
affected when removing cilia in the dorsal diencephalic cilia,
as the maximum velocity dropped from 27.7 \textmu m/s to 8.0 \textmu m/s.
\emph{In vivo}, these cilia generate the strongest flow at
this developmental stage of the fish~\cite{Olstad2019CiliaryDevelopment}.
In contrast, removing telencephalic and ventral diencephalic had minimal
impact and resulted in maximum velocities of 27.7 \textmu m/s or 27.2 \textmu m/s, respectively.
Changes in the transport dynamics of Dendra2 were most prominent when removing the dorsal cilia
in the diencephalic ventricle. All ROIs (except ROI 1, where we impose the photoconversion curve)
exhibit significant changes in the dynamics of $\cbar$
(\Cref{fig:figure5_compare_cilia_modifications}b).
Compared to the baseline model results, we observed a slower increase and lower final concentration
in the middle and anterior ventricles (ROIs 2--5),
and more rapid transport to the posterior ventricle (ROI 6).
On the other hand, removal of telencephalic cilia did not alter the evolution of the
mean concentration $\cbar$ with time (\Cref{fig:figure5_compare_cilia_modifications}b).
Finally, without ventral cilia in the diencephalic ventricle,
more solute has spread towards the anterior ventricle (ROI 5) and
less solute has spread towards the posterior ventricle (ROI 6).
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate ciliated regions where loss of cilia motility is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} Mean concentration $\cbar(t)$ for the six regions of interest in the three 
    loss-of-cilia-motility scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:figure5_compare_cilia_modifications}
\end{figure}

\subsection*{Modification of ventricles geometry impacts protein distribution}
Variations in ventricular morphology impacts the CSF flow patterns.
To assess the impact of ventricular morphologies on solute distribution,
we used four alternative brain ventricles meshes:
one where the connection between the anterior and middle ventricles is constricted;
one where the connection between the middle and posterior ventricles is constricted;
one where the middle ventricle is shrunk in the lateral direction;
and the combination of the above three modifications (\Cref{fig:figure6_sim_results_compare_geometries}a).
We simulate CSF flow in the modified geometries with the baseline flow model
and use the resulting velocity field to simulate transport of Dendra2
(\Cref{fig:figure6_sim_results_compare_geometries}b).

Constricting the connection between the anterior and middle ventricle leads to
slower transport to the anterior ventricle (ROI 5),
and a slightly higher mean concentration in ROIs 2--4 and 6, compared to the original geometry.
On the other hand, constricting the connection between the middle and the posterior ventricle
leads to less transport into the posterior ventricle (ROI 6),
and slightly more transport towards the anterior ventricle (ROI 5).
Faster transport to the anterior ventricle (ROI 5) and slower transport to the
posterior ventricle (ROI 6) is also the case for the geometry with a shrunk middle ventricle.
Additionally, for this configuration the transport upstream the flow fields towards the
ventral-posterior region of the middle ventricle (ROI 3) is slowed down.
For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have lower concentration.
Altogether our results identify that the geometry and width of the ventricular ducts matter
for solute transport and may explain why we observed larger differences
in the cilia-deficient model than in the experimental data in
\Cref{fig:figure4_compare_control_mutant}, which have different duct diameters.

\lyng{HH: Quantify differences in terms of percentage changes in the
final mean concentration values, maybe with a bar plot?}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{Caption. \mk{MK: How do the different geoemtries compare in terms of volume,
    areas where cilia are active?} \lyng{HH: Good question, will add this information.}}
    \label{fig:figure6_sim_results_compare_geometries}
\end{figure}

@@

\subsection*{Impact of cilia \fixme{properties} and ventricular geometry}
To gain further insights into how CSF distribution depends on cilia \fixme{properties (@Nathalie: do we really
investigate properties? Rather location or population?)} and geometries,
we manipulated certain parameters in our models. First, we removed cilia in specific regions of
our model (\Cref{fig:figure5_compare_cilia_modifications}). We identified that the cilia in the
dorsal middle ventricle contribute the most to solute distribution, which aligns well with the fact
that these cilia generate the strongest flow field \emph{in vivo} and in our model.
In contrast, removing the ventral cilia in the middle ventricle did not greatly impact flow dynamics.
However, it lead to reduced transport towards the posterior ventricle,
and increased transport towards the anterior ventricle, as the ventral cilia help induce
flow towards the posterior ventricles. Removing the cilia in the anterior ventricle did not affect the transport.

>>>>

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask: what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}a). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the \fixme{different-characteristics}
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}b). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}c, d). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site.
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of
    Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{b--d.} Slice of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for eight time instants covering the
    time range from 45 seconds to 790 seconds, corresponding to 100 to 1750 cardiac cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    The colorbar is scaled so the maximum is the maximum concentration of the Dendra2 
    condentration field slice at the last time instant.
    \textbf{e--g.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

To quantify the transport of solute, we plot the profile of the mean
concentrations $\cbar_i(t)$ as functions of time
(\Cref{fig:fig2}e--g) and
the times $\hat{t}$ when the mean concentrations $\cbar_i$ first exceed
a threshold value $\hat{c}$ (\Cref{fig:fig2}h)
in 6 regions of interest (ROIs) (\Cref{fig:fig2}a).
The dynamics at the site of photoconversion (ROI 1)
is logarithmic and follows from imposing \eqref{eq:photoconversion_curve}.
For ROIs 2--6, we observe faster kinetics with higher diffusion coefficients
(\Cref{fig:fig2}e--g),
which results in smaller times-to-threshold
(\Cref{fig:fig2}h).
For all three diffusion coefficients, the dynamics of ROI 5
is significantly delayed compared to ROI 2--4,
and ROI 6 is the slowest. Notably, for $D_1$ the mean concentration
in ROI 6 is close to zero at the end of simulation.
To quantify the balance of advection versus diffusion as transport mehcanisms,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$,
$\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the
dominant transport mechanism on a \fixme{global} scale.

\subsection*{Validation of the simulations with experimental data}
To validate our computational model, we performed photoconversion experiments
in transgenic zebrafish larvae expressing a secreted Dendra2 in their brain ventricles.
Our photoconversion protocol consisted of the acquisition of a baseline,
before localized exposure to a UV laser (at the ROI 1 location)
that converts Dendra2 from a green to a red emitting fluorescent protein. 
To quantify the transport of photoconverted proteins, we measured and
calculated the change in fluorescence intensity $\Delta F = (F(t)-F_0)/F_0$
over time (\Cref{fig:fig3}a,
$F_0$ being the average fluorescence at baseline).

We obtained very similar kinetics curves (\Cref{fig:fig3}a, b)
and time-to-threshold values (\Cref{fig:fig3}c, d) when comparing our
experiments (\Cref{fig:fig3}a, c) with
simulation (\Cref{fig:fig3}b, d).
Importantly, since the initial condition of the simulation is a mean concentration of zero,
the mean concentration $\cbar$ can be interpreted as the change in mean concentration and
is thus a similar metric to $\Delta F$.
The final values of $\cbar$ were 0.76, 0.72, 0.82, 0.56 and 0.35 respectively,
compared to 0.63, 0.71, 0.67, 0.29 and 0.18 for $\Delta F$.
Notwithstanding similarities, we observed small differences between the experiments and simulation.
For instance, transport dynamics in ROIs 2--6 were faster in the simulation and reached higher final values.
Moreover, the $\cbar$ profiles of ROIs 3--4 were opposite order in
the simulations as compared to the experiments.
Despite observing small differences, the \emph{in vivo} experiments and
simulations were similar enough to convincingly validate the model.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F$ over time
    in each region of interest (ROI)
    observed during experiments of photoconversion of Dendra2 proteins in ROI 1.
    \textbf{b.} Mean concentration $\cbar$ as function of time in each ROI in a
    simulation of Dendra2 protein transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F$ reaches a threshold value of 0.25 in
    one representative zebrafish larva.
    Time to $\cbar$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    We have excluded the posterior-most regions of the geometries,
    because the values of $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions.}
    \label{fig:fig3}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.7637806  0.72025859 0.81903216 0.56204164 0.34926411

\subsection*{Absence of ciliary motion affects local protein distribution}
Since CSF flow is highly disturbed in the absence of motile cilia, we
want to assess the impact of cilia motion on solute distribution. To this end,
we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant zebrafish (\emph{schmalhans} (\emph{smh})).
By plotting the means of $\Delta F$ (\Cref{fig:fig4}a)
and the time to threshold
across the control and mutant cohorts (\Cref{fig:fig4}b),
we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in $\sim 45\%$ (ROI 2) and $\sim 25\%$ (ROI 4)
higher mean times-to-threshold in the \emph{smh}
mutants compared to controls (\Cref{fig:fig4}b).

Next, to identify whether our model would recapitulate these findings,
we simulated protein transport by using the only-cardiac/no-cilia flow model 
and compared it with the baseline model
(both cilia and cardiac motion) (\Cref{fig:fig4}c).
Similar to our experiments, we observed slower dynamics in simulations for ROIs 2 and 4.
Interestingly, the difference was even greater for the simulations,
with the time-to-threshold in ROI 2 being 90\% higher in absence of cilia-mediated flow.
Moreover, ROI 5 exhibits very different $\cbar$ profiles for the two
configurations in our computational model,
but not in our experimental data.
Altogether, our findings identify that diffusion plays an important
role for solute movement to regions upstream in the cilia flow fields or to distant ventricles.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F$ over
    time in each region of interest (ROI) for control (black lines) and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets, and $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%.
    \textbf{c.} Mean concentration $\cbar$ as function of time in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

\subsection*{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we simulate transport
upon region-specific cilia paralysis by modifying the boundary conditions
of the CSF flow model. Three separate cilia-covered regions were considered separately:
the telencephalic (anterior) ventricle, the dorsal part and the ventral part
of the diencephalic (middle) ventricle (\Cref{fig:fig5}a).
The cilia tangential traction boundary condition is for each case
changed to a free-slip boundary condition in paralyzed region,
prescribing zero tangential forces to
represent absence of cilia motion.
We simulate CSF flow modeled by the Stokes equations with the
modified boundary conditions, and use the resulting
velocity fields in transport simulations.

We identified that the velocities were most strongly
affected when removing cilia in the dorsal diencephalic cilia,
as the maximum velocity dropped from 27.7 \textmu m/s to 8.0 \textmu m/s.
\emph{In vivo}, these cilia generate the strongest flow at
this developmental stage of the fish~\cite{Olstad2019CiliaryDevelopment}.
In contrast, removing telencephalic and ventral diencephalic had minimal
impact and resulted in maximum velocities of 27.7 \textmu m/s or 27.2 \textmu m/s, respectively.
Changes in the transport dynamics of Dendra2 were most prominent when removing the dorsal cilia
in the diencephalic ventricle. All ROIs (except ROI 1, where we impose the photoconversion curve)
exhibit significant changes in the dynamics of $\cbar$
(\Cref{fig:fig5}b).
Compared to the baseline model results, we observed a slower increase and lower final concentration
in the middle and anterior ventricles (ROIs 2--5),
and more rapid transport to the posterior ventricle (ROI 6).
On the other hand, removal of telencephalic cilia did not alter the evolution of the
mean concentration $\cbar$ with time (\Cref{fig:fig5}b).
Finally, without ventral cilia in the diencephalic ventricle,
more solute has spread towards the anterior ventricle (ROI 5) and
less solute has spread towards the posterior ventricle (ROI 6).
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate ciliated regions where loss of cilia motility is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} Mean concentration $\cbar(t)$ for the six regions of interest in the three 
    loss-of-cilia-motility scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:fig5}
\end{figure}

\subsection*{Modification of ventricles geometry impacts protein distribution}
Variations in ventricular morphology impacts the CSF flow patterns.
To assess the impact of ventricular morphologies on solute distribution,
we used four alternative brain ventricles meshes:
one where the connection between the anterior and middle ventricles is constricted;
one where the connection between the middle and posterior ventricles is constricted;
one where the middle ventricle is shrunk in the lateral direction;
and the combination of the above three modifications (\Cref{fig:fig6}a).
We simulate CSF flow in the modified geometries with the baseline flow model
and use the resulting velocity field to simulate transport of Dendra2
(\Cref{fig:fig6}b).

Constricting the connection between the anterior and middle ventricle leads to
slower transport to the anterior ventricle (ROI 5),
and a slightly higher mean concentration in ROIs 2--4 and 6, compared to the original geometry.
On the other hand, constricting the connection between the middle and the posterior ventricle
leads to less transport into the posterior ventricle (ROI 6),
and slightly more transport towards the anterior ventricle (ROI 5).
Faster transport to the anterior ventricle (ROI 5) and slower transport to the
posterior ventricle (ROI 6) is also the case for the geometry with a shrunk middle ventricle.
Additionally, for this configuration the transport upstream the flow fields towards the
ventral-posterior region of the middle ventricle (ROI 3) is slowed down.
For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have lower concentration.
Altogether our results identify that the geometry and width of the ventricular ducts matter
for solute transport and may explain why we observed larger differences
in the cilia-deficient model than in the experimental data in
\Cref{fig:fig4}, which have different duct diameters.

\lyng{HH: Quantify differences in terms of percentage changes in the
final mean concentration values, maybe with a bar plot?}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{Caption. \mk{MK: How do the different geoemtries compare in terms of volume,
    areas where cilia are active?} \lyng{HH: Good question, will add this information.}}
    \label{fig:fig6}
\end{figure}

@@

\subsection*{Impact of cilia \fixme{properties} and ventricular geometry}
To gain further insights into how CSF distribution depends on cilia \fixme{properties (@Nathalie: do we really
investigate properties? Rather location or population?)} and geometries,
we manipulated certain parameters in our models. First, we removed cilia in specific regions of
our model (\Cref{fig:fig5}). We identified that the cilia in the
dorsal middle ventricle contribute the most to solute distribution, which aligns well with the fact
that these cilia generate the strongest flow field \emph{in vivo} and in our model.
In contrast, removing the ventral cilia in the middle ventricle did not greatly impact flow dynamics.
However, it lead to reduced transport towards the posterior ventricle,
and increased transport towards the anterior ventricle, as the ventral cilia help induce
flow towards the posterior ventricles. Removing the cilia in the anterior ventricle did not affect the transport.
